---
title: "Homework 1 - P8105/Data Science I"
author: "Kevin S.W."
date: "9/12/2019"
output: html_document
---

```{r setup, include = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)
library(tidyverse)
library(dplyr)
```

# **Introduction**
This is an R-Markdown document for P8105 Homework 1. Tidyverse and dplyr was loaded through a chunk with `include = FALSE, message = FALSE, warning = FALSE`. This can be evaluated by accessing the actual .rmd file.

# Problem 1
## *Problem 1.1*
Making a data frame with:

* random sample; size 8, type = standard Normal distribution
* logical vector indicating whether elements of the sample are greater than 0
* Character vector; length = 8
* factor vector; length = 8, with 3 different factor “levels”
* Taking the mean of each variable in your dataframe. What works and what doesn’t?

```{r Problem_1}
## Seed used for reproducibility; remove for "true" function.
set.seed(2)

## Creates a data-frame that is composed of vectors below
prob1_df <- tibble(
  sample = rnorm(8),                              ## Vector for random normal-distributed sample of size 8
  vec_logic = sample > 0,                         ## Logical vector that evaluates if elements in sample > 0
  vec_char = c("GitHub", "is", "super",           ## Character vector with 8 variables ("length == 8")
               "interesting", "and", "awesome", 
               "I", "can't even..."),
  vec_fac = factor(c("corn", "not corn", "corn",  ## Factor vector size 8 with levels: "corn", 
              "maybe corn", "corn", "maybe corn", ## "maybe corn", "not corn"
              "corn", "not corn"))
)

## pull() function pulls the vectors inside prob1_df and turns it into usable variables outside the df
## Using prob1_df$vector was considered bad syntax and was not utilized
var_sample <- pull(prob1_df, sample)
var_vec_logic <- pull(prob1_df, vec_logic)
var_vec_char <- pull(prob1_df, vec_char)
var_vec_fac <- pull(prob1_df, vec_fac)

## Takes the mean for each vectors
mean(var_sample)
mean(var_vec_logic)
mean(var_vec_char)
mean(var_vec_fac)

```

Taking the mean of sample seems to work as expected as it is a continuous variable. Logical vector surprisingly spits out a mean of `r mean(var_vec_logic)`. Looking into `prob1_df` seems to indicate that `FALSE` = 0 while `TRUE` = 1, which leads to `r length(which(var_vec_logic == TRUE))`/`r length(var_vec_logic)` TRUE, thus the `r mean(var_vec_logic)`.

Mean for `var_vec_char` however, does not work. This is likely because characters are not read as "numbers" and therefore cannot be mathematically manipulated. 

## Problem 1.2
Write a code chunk that applies the as.numeric function to the logical, character, and factor variables (please show this chunk but not the output). What happens, and why? Does this help explain what happens when you try to take the mean?


## Problem 1.3
In a second code chunk:

convert the logical vector to numeric, and multiply the random sample by the result
convert the logical vector to a factor, and multiply the random sample by the result
convert the logical vector to a factor and then convert the result to numeric, and multiply the random sample by the result

```{r Problem_1c}

```

# Problem 2
Create a data frame comprised of:

* x: a random sample of size 500 from a standard Normal distribution
* y: a random sample of size 500 from a standard Normal distribution
* A logical vector indicating whether x + y > 1
* A numeric vector created by coercing the above logical vector
* A factor vector created by coercing the above logical vector
* Write a short description of your vector using inline R code, including: * the size of the dataset (using nrow and ncol) * the mean, median, and standard deviation of x * the proportion of cases for which x + y > 1

Make a scatterplot of y vs x; color points using the logical variable (adding color = ... inside of aes in your ggplot code should help). Make a second and third scatterplot that color points using the numeric and factor variables, respectively, and comment on the color scales.

Export your first scatterplot to your project directory using ggsave