---
title: "Homework 1 - P8105/Data Science I"
author: "Kevin S.W."
date: "9/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
```

# **Introduction**
This is a Rmarkdown document for P8105 Homework 1. Tidyverse and dplyr was loaded through a chunk with `include = FALSE`. This can be evaluated by accessing the actual .rmd file.

# Problem 1
## *Problem 1.1*
Making a data frame with:

* random sample; size 8, type = standard Normal distribution
* logical vector indicating whether elements of the sample are greater than 0
* Character vector; length = 8
* factor vector; length = 8, with 3 different factor “levels”
* Taking the mean of each variable in your dataframe. What works and what doesn’t?

```{r Problem 1, eval = TRUE}

prob1_df <- tibble(
  sample = rnorm(8),
  vec_logic = sample > 0,
  vec_char = "GitHub",
)

var_sample = pull(prob1_df, sample)
var_vec_logic = pull(prob1_df, vec_logic)
var_vec_char = pull(prob1_df, vec_char)

mean(c(var_sample, var_vec_logic, var_vec_char))
```

## Problem 1.2
Write a code chunk that applies the as.numeric function to the logical, character, and factor variables (please show this chunk but not the output). What happens, and why? Does this help explain what happens when you try to take the mean?

In a second code chunk:

convert the logical vector to numeric, and multiply the random sample by the result
convert the logical vector to a factor, and multiply the random sample by the result
convert the logical vector to a factor and then convert the result to numeric, and multiply the random sample by the result


```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
